---
title: "Regression models class project"
author: "Yevgen Yampolskiy"
date: "01/24/2015"
output: html_document
---

## Result summary
We observe significant difference in MPG for automatic and manual transmission. 
Most likely reason is that manual transmission weights less, 
and it is the weight impacts MPG.

Suggested model for MPG is `mpg ~ wt + cyl*hp`.

# Comparison of manual and automatic transmission for MPG
Just by looking at the box plots of MPG by transmission type we see that 
cars with manual transmission have higher MPG. T-test confirms that
the difference in mean MPG is significant. However it is more natural
to expect that weight will impact MPG a lot, and indeed we can see 
very high correlation between MPG and weight, and not so high correlation
between MPG and transimission type. If we plot MPG by weight and add 
coloring based on transmission type then we can obseved that cars
with manual transmission weight less which explains observed difference 
in MPG by transmission type. 

```{r}
data(mtcars)
boxplot(mpg ~ am, data = mtcars, xlab="Transmission type", ylab="MPG")
t.test(subset(mtcars, am==0, select=mpg), subset(mtcars, select=mpg, am==1))

# print correlations of MPG with transmission type and weight
c(cor(mtcars$mpg, mtcars$am), cor(mtcars$mpg, mtcars$wt)) 

#We can see that cars with manual transmission are more lightweight
library(ggplot2)
qplot(wt, mpg, data = mtcars, col = am)
```

# Choosing the model
If we feed all available variables into the model and
let `R` choose the best one using `step` function then resulting model
will be model `mpg ~ wt + qsec + am`. What I do not like about 
this model is that `qsec` is a derived variable from engine 
characteristic and weight, and `am` is likely related to weight.
Another issue is that using 3 variables on data set with 32 observations
could result into overfitting. 

If we try to model MPG using only two variables (one of them must be 
weight) then by looking at the adjusted `R^2` values the most attractive
models are `mpg ~ wt + cyl`, `mpg ~ wt + hp`, `mpg ~ wt + qsec`.
I like `cyl` variable in the model as it is directly related
to car physical characteristic. There is a high correlation between
`cyl` and `hp`. Adding interactions between
`cyl` and `hp` resulted into the model with the smallest `R^2`:
`mpg ~ wt + cyl * hp`.

Residual plot does not show anything bad about this model. Note
that if we compare `mpg ~ wt + qsec + am` and `mpg ~ wt + cyl * hp`
models than we reduced both residual error and number of degrees
of freedom. 

Here is supporting `R` code:
```{r}
data(mtcars)
lm1 <- lm(mpg ~ ., data = mtcars)
# Let R pick a model
step(lm1)
# Model choosen: mpg ~ wt + qsec + am
summary(lm(formula = mpg ~ wt + qsec + am, data = mtcars))
# Suggested model
model <- lm(formula = mpg ~ wt + cyl * hp, data = mtcars)
summary(model)
plot(resid(model))
abline(a=0, b=0)
```

